\normalfont\spacedallcaps{Hazardous asteroids forecast via Markov random fields}

### 2021-11-15 Manuscript/manuscript.tex
PUB

In Fig. \ref{FADM},\ref{FAMD_Quantitative variables} and \ref{FAMD_Individuals_(c)} are reported the main achivements of FADM (performed with FactoMineR package \cite{le2008factominer}): in particular from the correlation circle in Fig. \ref{FAMD_Quantitative variables} we can recognize different correlations that are given by the celestial mechanics law: for instance see that there is strong correlation of the mean motion with the semi-major axis. This is due to the Kepler law discussed in Appendix A. Furthermore we see that that the mean motion is correctly almost independent with respect to the the diameter (max or min) of  the asteroid. As explained in Appendix A this is an another result of Newton gravitation theory for a two body interaction: the mean motion of an asteroid, as long as the the mass of it is much lower with respect to the sun, is independent from its mass. On the other side the fact that relative velocity is correlated with the diameter is not surprising since here there is a three body dynamics (the Sun, the Earth and the Asteroid).

>>>>

In Fig. \ref{FADM},\ref{FAMD_Quantitative variables} and \ref{FAMD_Individuals_(c)} are reported the main achivements of FADM (performed with FactoMineR package \cite{le2008factominer}): in particular from the correlation circle in Fig. \ref{FAMD_Quantitative variables} we can recognize the different correlations that are given by the celestial mechanics laws: for instance see that there is strong correlation of the mean motion with the semi-major axis. This is due to the Kepler law discussed in Appendix A. Furthermore we see that that the mean motion is correctly almost independent with respect to the the diameter (max or min) of  the asteroid. As explained in Appendix A this is an another result of Newton gravitation theory for a two body interaction: the mean motion of an asteroid, as long as the the mass of it is much lower with respect to the sun, is independent from its mass. On the other side the fact that relative velocity is correlated with the diameter is not surprising since here there is a three body dynamics (the Sun, the Earth and the Asteroid). Beside this inspection a further analysis based on the concept of mutual information (summarized in the previous chapter) was performed: its result is reported in Fig. \ref{Mutual_information}. This figure summarize the ranking of the features considered in the dataset for the dangerousness classification of the asteroids according to the following expression \cite{kratzer2018varrank}

\begin{equation}
g(\alpha,\textbf{C},\textbf{S},f_{i})=MI(f_{i};\textbf{C})-\sum_{f_{s}\in S}\alpha(f_{i},f_{s},\textbf{C},\textbf{S})MI(f_{i};f_{s})
\end{equation}
where the first term $MI(f_{i};\textbf{C})$ is called relevance and measures the Mutual Information between the interesting feature set $\textbf{C}$ (only Hazardous in our case) and the analysed one $f_{i}$; the third term $MI(f_{i};f_{s})$ is called redundancy and measures the MI between the analysed feature and a chosen set of them. Finally the $\alpha(f_{i},f_{s},\textbf{C},\textbf{S})$ is a normalization function and in our case was set to \cite{kratzer2018varrank}

\begin{equation}
\alpha(f_{i},f_{s},\textbf{C},\textbf{S})=\dfrac{1}{|\textbf{S}|}
\end{equation}

following the Peng. et al approach \cite{peng2005feature}. We see that the first placein the mutual information ranking, looking to the diagonal element,  is taken by minimum orbit intersection: this is correct since this is one parameter used by NASA for deciding if an asteroid is hazardous or not (see Appendix A). The second place is occupied by Epoch date close approach. The third place is the eccentricity value: this parameter is entangled with the minimum orbit intersection and thus it is reasonable that it is important. Then we have two parameter that are related to the dimension of the asteroid: this is meaningful since if the asteroid has a too reduced volume it will be destroyed by the Earth atmosphere. On the other side the other parameters seems to have a too low MI for being interesting in this preliminary analysis. The results obtained for FAMD and MI will provide a useful path-guide, together with the celestial mechanics theory, for the interpretation of the results obtained in the following section
